How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)

### 2015-08-24 Manuscript.tex
Updates from ShareLaTeX

\section{Preliminaries}
\subsection{From Prediction to Planning}

@@

Note that predictions  about changes that  fall within the space of the training+test data, will be at least
as reliable as the performance value found in the data mining experiment.
Given this, one thing  can be asserted about predictions on changed examples:
\bi
\item Predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
\ei  
Accordingly, all the methods of this paper make {\em trust-increasing} recommendations. That is, they push change examples {\em towards} the
training examples.  To see how this works, 
 \fig{howxy} is from the {\em ivy} data
set, which is one of the Jureczko data set. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
changed  examples displaced after applying one of our planning methods.
To generate that figure, we placed all of 1+2+3 into the same data set then computed the first
two  components (using PCA). Note that the  the   changed examples
cases  (shown in green)  fall closer to the training cases (shown in gray) than
the test cases (shown in red).  That is, if we believe
that the gray data is  suitable for assessing
the test cases (as indicated by the {\em ivy} results from \fig{j}), then it follows
that we believe even more in the predictions
generated from the displaced data).

@@

That said, the above comes with two important caveats:
\bi
\item 
We   strongly recommend that predictors are assessed prior to planning. That
issue is explored further in the next section. 
\item
Where possible, planners should be assessed via some external
oracle that can accurately assess new examples. For an example of that kind of analysis,
see our Validity section, later in this paper.
\ei

  \section{Test Domains}
To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
  See github.com/ai-se/XTREES\#data for full access to this data.

@@

 It proved to be  more complicated to commission the Jureczko data sets for this study.
 For that data, we found:
\bi
\item
 Quality predictors built from this data are far from perfect;
\item 
But several of these predictors could
be salvaged using the techniques discussed in this section.
\ei
 \fig{j} shows our preliminary studies with the Jureczko   data.
Given access to $V$ released
versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as
shown in \fig{j}, this means that we are training on hundreds to thousands
of classes and testing on smaller test suites).
Note the three bottom \colorbox{lavenderpink}{highlighted} rows marked with $\times$: these contain predominately
defective classes (two-thirds, or more).  In such data sets, it is hard to distinguish
good from bad (since there are so many bad examples).

The  Jureczko data uses non-numeric discrete classes (``defective'' or ``not'').
For such data, quality predictor   is be measured using:
\bi
\item Probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
the test data detected
by the {\em predictor}.
\item Probability of false alarm (a.k.a. ``pf''): the percent of non-fault
classes that are {\em predicted} to be defective.
\ei 
As a preliminary study, we split the Jureczko  data   into equal sized train:test groups.
Random Forests (again, from the SciKit learn kit~\cite{Pedregosa2012}) were
built from the training data, then applied to the test data.
The ``untuned'' columns of \fig{j} shows those results.
If we define ``good'' to mean $\mathit{pd}>60 \wedge \mathit{pf} < 40$\%,
then only two of our data sets ({\em ivy,ant}) are ``good'' enough for this study.
Note that, as might have been expected, none of the three bottom rows of \fig{j} were ``good''.

Fortunately,
the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets. Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e. an over-sampling of minority-class examples. Also, Fu et al.~\cite{fu:fse16} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
for tree generation, etc. The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these techniques. For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%. However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.

@@

	For  continuous and discrete values,
the {\em variability} can be measured using standard deviation $\sigma$ or entropy $e$.
Note that \mbox{$\sigma = \sqrt{\sum^n_{i=1} \left( \frac{x_i - \bar{x})^2}{n-1}\right)}$}
where $\bar{x}$ is the mean of  numeric features $x_1,x_2,..x_n$.
Also \mbox{$e = - \sum^n_{i=1} p_i \mathit{log}_2(p_i)$} 
for  $n$ discrete values occurring at frequency
$f_1,f_2,.. f_n$. That is, \mbox{$N = \sum^n_{i=1} f_i$} and $p_i = f_i/N$.

@@

				WHERE  divides data into  groups of size $\alpha=\sqrt{N}$ using 
		Using this measure, WHERE runs as follows:
		\begin{enumerate}[leftmargin=3mm]
			\item Find   two   distance cases,  $X,Y$
			by picking any case $W$ at random, then setting $X$ to its most
			distant case, then setting $Y$ to the case most distant from
			$X$
			(which requires only $O(2N)$ comparisons).
			\item Project each case $Z$
			to position $x$ along a    lines that  runs from $X$ to $Y$, If $a,b$ re distances  $Z$ to $X,Y$  then  $x = (a^2+c^2 - b^2)/(2ac)$.
			\item Split the data at the median $x$ value of all cases.
			\item For   splits larger than  $\alpha=\sqrt{N}$, call this algorithm recursively  from step1.
		\end{enumerate}
 In terms of related work,
	  the above is similar in approach to Boley's PDDP algorithm~\cite{boley98}, but PDDP requires an $O(N^2)$ calculation
	  at each recursive level to find the PCA principle component. Our method, on the other hand,
	  performs the same task with only $O(2N)$ distance calculations 	using the 
	  FASTMAP heuristic~\cite{Faloutsos1995} shown in step1. Platt~\cite{platt05} notes that FASTMAP is a  Nystr\"om approximation to the first component of PCA.

@@

	Find a split in the values of the independent features that most reduces the variability
  of the dependent feature\footnote{}. The data is then divided on that feature's split and 
  the tree learner recurses on each split (stopping when any split is smaller than $\alpha=\sqrt{N}$).

@@

\section{Four Planning Methods}
In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
algorithm should be baselined against simpler alternatives. Accordingly,
this section described XTREES (which we call Method4) against three  
alternatives: 
  \bi
  \item  
  Method1, Method2 and Method3 use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	\item
 Method4 uses the decision tree learner of \fig{where}.D. 
  \ei

  Note that all our methods are {\em trust-increasing} in the manner described in \tion{trust}.

The following descriptions of the methods adopts the following convention. All variables
  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma$.
  In this paper, we show our current settings to these variables produces useful
  results. In other work~\cite{krall14}, we are exploring tuning methods to automatically
  find better settings but  we have nothing definitive yet to report
  on auto-tuning planners.

@@

 {\em Details:} Method1 clusters project data by reflecting   on the independent variables, then
  reporting the delta between the cluster centroids. The advantage of Method1 is its
  simplicity. However, as shown by our evaluation strategy, it is not an useful
  method for generating effective plans.

After clustering some training data using the WHERE algorithm of  \fig{where}.C, this method:
replaces all clusters with a  centroid $C_i$ computed from the mean/mode value of each
continuous/discrete feature. After that, it
finds the closest centroid $C_j$ that has a better
performance score. For defect data, ``better'' means fewer defects while for the config data,
``better'' means faster runtimes.
Next, this method caches the  delta in the independent features between $C_i$ and $C_j$. For continuous
features, this delta is $C_j - C_i$. For discrete values, this delta is the value of that feature
in $C_j$. 
Finally, for every test case, Method1 use the distance measure $d$ shown in \fig{where}.B to find
the nearest centroid $C_i$.  It then proposes a plan for improving that test case
that is the conjunction of all the deltas between $C_i$ and $C_j$.

\subsection{Method2 = CD+FS =  Method1  + \\Feature Selection }

>>>>

\section{Preliminaries}\label{sect:prelim}
\subsection{From Prediction to Planning}

@@

Note that predictions  about changes that  fall within the space of the training+test data, will be at least
as reliable as the performance value found in the data mining experiment.
Given this, one thing  can be asserted about predictions on changed examples:
\bi
\item Predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
\ei  
Accordingly, all the methods of this paper make {\em trust-increasing} recommendations. That is, they push change examples {\em towards} the
training examples.  To see how this works, 
 \fig{howxy} is from the {\em ivy} data
set, which is one of the Jureczko data sets explored in this paper. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
changed  examples displaced after applying one of our planning methods.
To generate that figure, we placed all of 1+2+3 into the same data set then computed the first
two  components (using PCA). Note that the  the   changed examples
cases  (shown in green)  fall closer to the training cases (shown in gray) than
the test cases (shown in red).  That is, if we believe
that the gray data is  suitable for assessing
the test cases (as indicated by the {\em ivy} results from \fig{j}), then it follows
that we believe even more in the predictions
generated from the displaced data).

@@

That said, the above comes with two important caveats:
\bi
\item 
We   strongly recommend that predictors are assessed prior to planning. That
issue is explored further in the next section. 
\item
Where possible, planners should be assessed via some external
oracle that can accurately assess new examples. For an example of that kind of analysis,
see the Validity section, later in this paper.
\ei

  \section{Test Data}
To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
  See github.com/ai-se/XTREES\#data for full access to this data.

@@

 It proved to be  more complicated to commission the Jureczko data sets for this study.
 For that data, we found that the
 quality predictors built from this data are far from perfect;
However, for some data sets, the  predictors could
be salvaged using the techniques discussed in this section.

 \fig{j} shows our preliminary studies with the Jureczko   data.
Given access to $V$ released
versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as
shown in \fig{j}, this means that we are training on hundreds to thousands
of classes and testing on smaller test suites).
Note the   \colorbox{lavenderpink}{three bottom rows}   marked with $\times$: these contain predominately
defective classes (two-thirds, or more).  In such data sets, it is hard to distinguish
good from bad (since there are so many bad examples).

The  Jureczko data uses non-numeric discrete classes (``defective'' or ``not'').
For such data, quality predictor   is be measured using:
\bi
\item Probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
the test data detected
by the {\em predictor}.
\item Probability of false alarm (a.k.a. ``pf''): the percent of non-fault
classes that are {\em predicted} to be defective.
\ei 
As a preliminary study, we split the Jureczko  data   into equal sized train:test groups.
Random Forests (again, from the SciKit learn kit~\cite{Pedregosa2012}) were
built from the training data, then applied to the test data.
The ``untuned'' columns of \fig{j} shows those results.
If we define ``good'' to mean $\mathit{pd}>60 \wedge \mathit{pf} < 40$\%,
then only two of our data sets ({\em ivy,ant}) are ``good'' enough for this study.
Note that, as might have been expected, none of the \colorbox{lavenderpink}{three bottom rows} of \fig{j} were ``good''.

Fortunately,
the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets. Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e. an over-sampling of minority-class examples. Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
for tree generation, etc. The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these techniques. For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%. However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.

@@

	For  continuous and discrete values,
the {\em variability} can be measured using standard deviation $\sigma$ or entropy $e$.
Note that \mbox{$\sigma = \sqrt{\sum^n_{i=1} \left( \frac{x_i - \bar{x})^2}{n-1}\right)}$}
where $\bar{x}$ is the mean of  numeric features $x_1,x_2,..x_n$.
Also \mbox{$e = - \sum^n_{i=1} p_i \mathit{log}_2(p_i)$} 
for  $n$ discrete values  at frequency
$f_1,f_2,.. f_n$ for \mbox{$N = \sum^n_{i=1} f_i$} and $p_i = f_i/N$.

@@

				WHERE  divides data into  groups of size $\alpha=\sqrt{N}$ using 
		Using this measure, WHERE runs as follows:
		\begin{enumerate}[leftmargin=3mm]
			\item Find   two   distance cases,  $X,Y$
			by picking any case $W$ at random, then setting $X$ to its most
			distant case, then setting $Y$ to the case most distant from
			$X$
			(which requires only $O(2N)$ comparisons).
			\item Project each case $Z$
			to position $x$ on a    lines running from $X$ to $Y$: if $a,b$  are distances  $Z$ to $X,Y$  then  $x = (a^2+c^2 - b^2)/(2ac)$.
			\item Split the data at the median $x$ value of all cases.
			\item For   splits larger than  $\alpha=\sqrt{N}$, recurse   from step1.
		\end{enumerate}
 In terms of related work,
	  the above is similar in approach to Boley's PDDP algorithm~\cite{boley98}, but PDDP requires an $O(N^2)$ calculation
	  at each recursive level to find the PCA principle component. Our method, on the other hand,
	  performs the same task with only $O(2N)$ distance calculations 	using the 
	  FASTMAP heuristic~\cite{Faloutsos1995} shown in step1. Platt~\cite{platt05} notes that FASTMAP is a  Nystr\"om approximation to the first component of PCA.

@@

	Find a split in the values of  independent features that most reduces the variability
  of the dependent feature (measured using \fig{where}.A). Divide data on that  feature's split and 
   for all splits bigger than $\alpha=\sqrt{N}$,  recurses on each split.

@@

\section{Four Planning Methods}
In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
algorithm should be baselined against simpler alternatives. Accordingly,
this section described XTREES (which we call Method4) against three  
alternatives: 
  \bi
  \item  
  Method1, Method2 and Method3 use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	\item
 Method4 uses the decision tree learner of \fig{where}.D. 
  \ei
  Note that all our methods have the two properties proposed in \tion{prelim}:
  \bi
  \item They find different plans in different parts of the data;
  \item They  are {\em trust-increasing} in the manner described in \tion{trust};
  i.e. they change changing examples such that they move {\em closer} to the training data.
  \ei
The following descriptions of the methods adopts the following convention. All variables
  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma$.
  In this paper, we show our current settings to these variables produces useful
  results. Elsewhere\cite{krall14,fu:ase15}, we are exploring tuning methods to 
  find better settings but  we have nothing definitive yet to report
  on auto-tuning planners.

@@

 {\em Details:} Method1 clusters project data by reflecting   on the independent variables, then
  reporting the delta between the cluster centroids. 
  After clustering   training data using the WHERE algorithm of  \fig{where}.C, Method1
replaces all clusters with a  centroid $C_i$ computed from the mean/mode value of each
continuous/discrete feature. After that, it
finds the closest centroid $C_j$ that has a better
performance score. For defect data, ``better'' means fewer defective examples while for the config data,
``better'' means lower median runtimes for the examples in that cluster.
Method1 then caches the  delta between the independent features between $C_i$ and $C_j$. For continuous
features, this delta is $C_j - C_i$. For discrete values, this delta is the value of that feature
in $C_j$. 
Finally, for every test case, Method1 use the distance measure $d$ shown in \fig{where}.B to find
the nearest centroid $C_i$.  It then proposes a plan for improving that test case
that is the conjunction of all the deltas between $C_i$ and $C_j$.

\subsection{Method2 = \\\CD+FS =  Method1  + Feature Selection }

### 2015-08-24 tex/prediction_stats.tex
Updates from ShareLaTeX

\begin{figure*}[bp]
	\small
	\begin{center}
		\begin{minipage}{.46\linewidth}
			\begin{tabular}{r@{~}|l@{~}|r@{~}|l@{~}|r@{~}|r@{~}|} \cline{2-6}
				& \multicolumn{5}{c|}{ }\\

@@

	\caption{Training and test {\em data set properties} for the Jureczko data sets,
		sorted in ascending order of \% defective examples.
		On the right-hand-side, we show the {\em results from learning}.
		Data is ``good'' if it has recall over 60\% and false alarm under 40\%
		(and note that, after tuning, there are more ``good'' than before).
		Data   marked with ``$\star$'' show large improvements in performance, after tuning.
		The data in  the  \colorbox{lavenderpink}{highlighted} rows, marked with ``$\times$'', are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
	}\label{fig:j}
\end{figure*}

>>>>

\begin{figure*}[!t]
	\small
	\begin{center}
		\begin{minipage}{.46\linewidth}
			\begin{tabular}{r@{~}|l@{~}|r@{~}|l@{~}|r@{~}|r@{~}|} \cline{2-6}
				& \multicolumn{5}{c|}{ }\\

@@

	\caption{Training and test {\em data set properties} for  Jureczko data ,
		sorted by \% defective examples.
		On the right-hand-side, we show the {\em results from learning}.
		Data is ``good'' if it has recall over 60\% and false alarm under 40\%
		(and note that, after tuning, there are more ``good'' than before).
		Data   marked with ``$\star$'' show large improvements in performance, after tuning
		(lower {\em pf} or higher {\em pd}).
		The data in  the  \colorbox{lavenderpink}{three bottom rows}, marked with ``$\times$'', are ``not good'' since their test suites  have so few non-defective examples (less than 5\% of the total sample) that it becomes harder to find better data towards which we can displace test data.
	}\label{fig:j}
\end{figure*}
